---
title: "My Fisheries Analysis"
author: "ESellers"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# attach libraries
library(tidyverse)
library(here)
library(janitor)
library(paletteer)
```

Read in noaa_landings.csv data (fixing the 'no data' values to proper NAs)
```{r read data}
us_landings <- read_csv(here("data/noaa_landings.csv"), 
                        na = "no data")
```

Clean up the dataset
- clean the field labels and in the newly labeled afs_name field convert values to lowercase
```{r clean the dataset}
salmon_clean <- us_landings %>% 
  clean_names() %>% 
  mutate(afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_num = parse_number(dollars_usd)) %>% 
  filter(str_detect(afs_name, pattern = "salmon")) %>% 
  separate(afs_name, into = c("salmon_group", "subgroup"), sep = ", ") %>% 
  drop_na(dollars_num)
```

sep - ", " captures the space along with the separator so that the space does not get copied over
NOTE: drop_na(dollars_num) deletes the entire row where dollars_num is NA

### Use a pivot table to create an Annual total value by salmon group

```{r}
salmon_us_annual <- salmon_clean %>% 
  group_by(year, subgroup) %>% 
  summarize(
    tot_value = sum(dollars_num, na.rm = TRUE)
  )
```

Activity: Create a line graph from salmon_us_annual with year on the x axis and tot_value on the y axis, that has a diff line for each subgroup

```{r create a line graph}
ggplot(data = salmon_us_annual, aes(x = year, y = tot_value, 
                                    group = subgroup)) + 
  geom_line(aes(color = subgroup)) + 
  theme_bw() +
  labs(x = "Year", 
       y = "US commercial salmon value (USD)", 
       title = "my awesome graph") + 
  scale_color_paletteer_d("ggsci::uniform_startrek")

# scale_color_paletteer_d for 'discrete'
# ggsci	uniform_startrek
```

















